---
title: "Homework 7: Strings + Logical Vectors (Graphical Perception)"
author: "MICR 475/575 Hadley Donaldson"
format:
  html:
    toc: true
    embed-resources: true
execute:
  warning: false
  message: false
editor: visual
---

## Instructions

In this homework you will practice:

-   importing a Google Forms CSV export
-   working with **strings**
-   building and using **logical vectors**
-   making data tidy with `pivot_longer()`
-   making a simple **box plot**

You will work with a Google Forms CSV export from our in-class graphical perception experiment.

------------------------------------------------------------------------

## Setup

```{r}
library(tidyverse)

csv_path <- "handb_ptest (3).csv"

```

------------------------------------------------------------------------

# Q1. Import the data

**Task:** Read the CSV into an object called `raw`, then preview it.

```{r}

raw <- read_csv(csv_path)

glimpse(raw)

```

------------------------------------------------------------------------

## Q2. Tidy the data (Method A: pivot_longer first, then rename question type variable)

In the raw Google Forms export, the response columns have long names that include the full question text. Your goal is a tidy data frame with one row per response, and columns:

-   `timestamp`
-   `chart_type` (values like `T1`, `T2`, ..., `T9`)
-   `percent` (numeric)

**Task A:** Start from `raw`, `pivot_longer()` everything *except* `timestamp`.

**Task B:** Convert the `percent` column to a number.

**Task C:** Use **string operations** (and at least one **logical vector**) to convert the messy `chart_type` column into clean labels.

> Hint 1: `parse_number(chart_type)` can pull the number out of the string.\
> Hint 2: A logical vector can help you keep only the question numbers you want.

**Task D:** Make a **box plot** of `percent` by `chart_type` (no log transform).

```{r}

tidy_a <- raw %>%
  rename(timestamp = Timestamp) %>%
  pivot_longer(
    cols = -timestamp,
    names_to = "chart_type",
    values_to = "percent"
  ) %>%
  mutate(
    keep_percent = str_detect(chart_type, "approximately")
  ) %>%
  filter(keep_percent) %>%
  mutate(
    percent = parse_number(percent),
    q_num = parse_number(chart_type),
    chart_type = if_else(is.na(q_num), "T1", paste0("T", q_num + 1))
  ) %>%
  select(timestamp, chart_type, percent)

tidy_a

ggplot(tidy_a, aes(x = chart_type, y = percent)) +
  geom_boxplot() +
  labs(
    title = "Percent estimates by chart type",
    x = "Chart type",
    y = "Percent"
  ) +
  theme_minimal()
```

------------------------------------------------------------------------

# Q3. Tidy the data (Method B: rename columns first, then pivot_longer)

This method uses the idea that (after `Timestamp`) the remaining columns are the 9 chart types, in order. The following steps provide one way this can be done - if you can get to the right answer using a different method, that's fine too!

**Task A:** Create a new data frame `df` that keeps `timestamp` plus the 9 response columns.

**Task B:** use the paste0() function to generate a vector of neat column headings 'T1', 'T2', etc. (keep `timestamp` as-is).

**Task C:** use the function set_names() with your new vector of column names to rename your columns

**Task D:** `pivot_longer()` to make a tidy data frame with `timestamp`, `chart_type`, `percent` (numeric).

**Task E:** Make the same **box plot** of `percent` by `chart_type`.

```{r}

df <- raw %>%
  select(Timestamp, contains("approximately"))

new_names <- c("timestamp", paste0("T", 1:9))

df <- df %>%
  set_names(new_names)

tidy_b <- df %>%
  pivot_longer(
    cols = -timestamp,
    names_to = "chart_type",
    values_to = "percent"
  ) %>%
  mutate(
    percent = parse_number(percent)
  )

tidy_b

ggplot(tidy_b, aes(x = chart_type, y = percent)) +
  geom_boxplot() +
  labs(
    title = "Percent estimates by chart type",
    x = "Chart type",
    y = "Percent"
  ) +
  theme_minimal()

```

------------------------------------------------------------------------

# Q4. Do the two methods agree?

**Task A:** Check that the boxplots from Q2 and Q3 are the same. It's ok if they aren't - just please comment on why they disagree

**Task B:** In 3–6 sentences, comment on what the box plots show. Does it give the same gist that we would expect from Heer and Bostock's results? If not, why not?

```{r}

arrange_a <- tidy_a %>%
  arrange(timestamp, chart_type, percent)

arrange_b <- tidy_b %>%
  arrange(timestamp, chart_type, percent)

identical(arrange_a, arrange_b)

#task b: the two methods agree because they both are isolating the same 9 percent response comlumns and reshape them into tidy format with one row per response. The box plots both show that these responses vary accross the chart types. This shows us that some visual formats are easier for people to judge more accurately than others. This also generally matches the main idea presented by Heer and Bostock, that graphical perception depends on the type of visuals used. Our results do not match perfectly because our class sample is a lot smaller and includes some missing responses and word responses ("about the same"). 
```
