---
title: "Homework 7: Strings + Logical Vectors (Graphical Perception)"
author: "MICR 475/575"
format:
  html:
    toc: true
    embed-resources: true
execute:
  warning: false
  message: false
editor: visual
---

## Instructions

In this homework you will practice:

-   importing a Google Forms CSV export
-   working with **strings**
-   building and using **logical vectors**
-   making data tidy with `pivot_longer()`
-   making a simple **box plot**

You will work with a Google Forms CSV export from our in-class graphical perception experiment.

------------------------------------------------------------------------

## Setup

```{r}
library(tidyverse)

csv_path <- "data/handb_ptest.csv"
```

------------------------------------------------------------------------

# Q1. Import the data

**Task:** Read the CSV into an object called `raw`, then preview it.

```{r}
# YOUR CODE HERE
raw <- read.csv("M:/Reproducible Data/Homework 7/r4ds_c10_to_16/data/handb_ptest.csv", na = c("N/A",""))

Header <- read_csv("M:/Reproducible Data/Homework 7/r4ds_c10_to_16/data/handb_ptest.csv", n_max = 0)

```

------------------------------------------------------------------------

## Q2. Tidy the data (Method A: pivot_longer first, then rename question type variable)

In the raw Google Forms export, the response columns have long names that include the full question text. Your goal is a tidy data frame with one row per response, and columns:

-   `timestamp`
-   `chart_type` (values like `T1`, `T2`, ..., `T9`)
-   `percent` (numeric)

**Task A:** Start from `raw`, `pivot_longer()` everything *except* `timestamp`.

**Task B:** Convert the `percent` column to a number.

**Task C:** Use **string operations** (and at least one **logical vector**) to convert the messy `chart_type` column into clean labels.

> Hint 1: `parse_number(chart_type)` can pull the number out of the string.\
> Hint 2: A logical vector can help you keep only the question numbers you want.

**Task D:** Make a **box plot** of `percent` by `chart_type` (no log transform).

```{r}
raw_long <- raw %>%
  mutate(response_id = row_number()) %>%       
  pivot_longer(
    cols = 2:19,                              
    names_to = "question",
    values_to = "score"
  )
raw_long <- raw_long %>%
  mutate(score = parse_number(as.character(score))) 

raw_long <- raw_long |> filter(!is.na(score))

raw_long<- raw_long |> mutate(question = str_replace(question, "The.smaller.is.approximately.___..of.the.larger", "T"))
ggplot(raw_long,
       aes(x = question, y = score)) +
  geom_boxplot()+
  labs(
    x = "Chart Type",
    y = "Percent",
    title = "Percent by Chart Type"
  )
```

------------------------------------------------------------------------

# Q3. Tidy the data (Method B: rename columns first, then pivot_longer)

This method uses the idea that (after `Timestamp`) the remaining columns are the 9 chart types, in order. The following steps provide one way this can be done - if you can get to the right answer using a different method, that's fine too!

**Task A:** Create a new data frame `df` that keeps `timestamp` plus the 9 response columns.

**Task B:** use the paste0() function to generate a vector of neat column headings 'T1', 'T2', etc. (keep `timestamp` as-is).

**Task C:** use the function set_names() with your new vector of column names to rename your columns

**Task D:** `pivot_longer()` to make a tidy data frame with `timestamp`, `chart_type`, `percent` (numeric).

**Task E:** Make the same **box plot** of `percent` by `chart_type`.

```{r}
df <- raw |>
  select(Timestamp, starts_with("the"))


my_names = c('timestep',paste0())

my_names = c('timestep',paste0('q',1:9))


colnames(df)<- my_names

df_long <- df |>
  pivot_longer(
    cols = q1:q9,
    names_to = "chart_type",
    values_to = "percent"
  ) |>
  mutate(percent = as.numeric(percent))

ggplot(df_long, aes(x = chart_type, y = percent)) +
  geom_boxplot() +
  labs(
    x = "Chart Type",
    y = "Percent",
    title = "Percent by Chart Type"
  )

```

------------------------------------------------------------------------

# Q4. Do the two methods agree?

**Task A:** Check that the boxplots from Q2 and Q3 are the same. It's ok if they aren't - just please comment on why they disagree

**Task B:** In 3–6 sentences, comment on what the box plots show. Does it give the same gist that we would expect from Heer and Bostock's results? If not, why not?

```{r}
#Task A: Both of the boxplots looks similar because the box plots comes from the same dataset and the data was tidyed in the same manner. The process of tidying may be differnet but at the end both data set was transformed to a longer version of the data and then ploted with the same parameters so they both look same.

#Task B: The box plots show the distribution of participant estimates for each chart type. Some chart types show wider spreads, indicating that people had more difficulty estimating the percentage accurately. Other chart types have tighter boxes, suggesting more consistent estimates among participants.Overall, the results generally align with the findings of Heer and Bostock, who showed that some visual encodings (such as position or aligned bars) allow people to judge quantities more accurately than others (such as angles or areas).
```
